\vspace{-2cm} Kronos \& Krios: Evidence for accretion of rocky planets in a comoving pair of G dwarfs \vspace{-0.5cm}

### 2017-06-23 paper/ms.tex
many tweaks; break page before supplement

% general description of the pair and measurements
\sunanalog\ and \bizarreone\ were identified as a candidate comoving star pair
in our recent search for comoving stars using the proper motions and parallaxes
from the {\it Tycho-Gaia Astrometric Solution} catalog (hereafter \tgas), a
component of the first data release of the astrometric space mission \gaia\cite{2016A&A...595A...2G}
(the astrometric measurements are listed in \tablename~\ref{tab:t1}).
The pair has also been previously recognized as a visual double star system
in Washington Double Star catalog\cite{2001AJ....122.3466M}.
The two stars have spectral types G0 and G2 both similar to the Sun (G2).
In a separate effort to study detailed chemical abundances of potential
planet-hosting stars, high resolution ($R=\lambda/\Delta\lambda\approx 70000$),
high signal-to-noise ratio (typically $>200$ per pixel in continuum) spectra of
both stars were obtained using the HIRES spectrograph on the Keck-I
telescope\cite{2016ApJS..225...32B}. 
The wavelength coverage is $5164$--$7799$~\AA.
The resulting measurements include elemental abundances for 15 chemical species
(C, N, O, Na, Mg, Al, Si, Ca, Ti, V, Cr, Mn, Fe, Ni, Y) as well as stellar
parameters and high precision ($\sigma\approx0.2$~\kms) radial velocities.

% likely origin given its proximity in phase space
% TODO reference dxdv figure here
The projected separation between the pair is 1.9 arcmin\ ($\approx 0.01$~pc),
and the 3D separation is $\approx 0.6$~pc.
Although selected based only on their astrometry, the two stars
have identical radial velocities within uncertainties (Table~\ref{tab:t1}),
confirming that they are truly comoving.
As the separation is smaller than the Jacobi radius (1.2~pc) in the solar
neighborhood for a 2~\msun\ binary system\cite{Jiang:2010aa}, \bizarreone\ and
\sunanalog\ are likely a bound wide binary system that formed coevally.

% coevality based on stellar parameters
% TODO reference supplement discussion on age here?
Their stellar parameters also indicate that they are coeval with an age of $\sim 4$~Gyr.
We use the distances (inferred from \gaia\ parallaxes), $V$-band magnitudes,
and $B-V$ colors to obtain bolometric luminosities of the two stars\cite{2003AJ....126..778V}.
We then combine the luminosities with $T_\mathrm{eff}$, \elemH{Fe}, \elemH{Si}
in order interpolate the age, mass, and radius of each star
in a grid of Yale-Yonsei model isochrones\cite{2013ApJ...776...87S}.
The best-fit isochrone ages of \sunanalog\ and \bizarreone\ are
$4.00_{-1.56}^{+1.51}$~Gyr and $4.28_{-1.03}^{+1.11}$~Gyr, respectively,
consistent with them being coeval (see Figure~\ref{fig:yyisofitting}).

% metallicity and abundance difference
However, one of the stars, \bizarreone\, is significantly more metal rich than
the other (by 0.2~dex $\approx 60\%$; \figname~\ref{fig:abundances}).
Moreover, not all elements are equally enhanced: the abundances of \bizarreone\
show selective under-enhancement in \elem{C}, \elem{N}, \elem{O}, \elem{Na},
and \elem{Mn} relative to \elem{Fe} and other refractory elements.
As expected from their measured metallicity difference, the ratio of data and
model between the two stars show significant residuals of almost all metal line
features, largely dominated by \elem{Fe}.
However, for lines of elements that are not as enhanced in \bizarreone\, the
residuals are much smaller in amplitude demonstrating the validity of the
measured abundance differences (see
Figure~\ref{fig:spec1} and \ref{fig:spec2}).

% novelty
The observed difference in metallicity and abundances are genuinely rare and
unexpected.
None of the other four twin-like ($\Delta T_\mathrm{eff} \lesssim 100$~K) wide
binary pairs examined using spectra of similar quality and the same
consistent measurement pipeline shows discrepancies in abundances between the
stars at this level\cite{2016ApJS..225...32B} (see Figure~\ref{fig:deltaXH}).
The maximum metallicity difference observed in the other pairs is $\Delta\feh =
0.03$~dex with $1\sigma$ statistical uncertainties in \feh\ of $0.01$~dex.
This is consistent with similar studies of metallicity and abundance
differences in wide binaries\cite{Gratton:2001aa,Desidera:2004aa}.
The difference of $\approx 0.2$~dex is also the largest among a handful of wide
binaries that do show metallicity and abundance differences with varying
significance\cite{Mack:2014aa,Mack:2016aa,Saffe:2015aa,
  Teske:2013aa,Teske:2015aa,Teske:2016aa,Teske:2016ab,Biazzo:2015aa,Ramirez:2015aa}
(see Figure~\ref{fig:relabun_tcrank}).

% high Li
Both stars have high surface \elem{Li} abundances considering
their ages of $\sim 4$~Gyr.
The surface lithium abundance in a sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning. %TODO: REFERENCE?
The \elem{Li} doublet at $6707.6$~\AA, which is usually too weak to be
detectable at this age even for high signal-to-noise spectra, is clearly
visible in the spectra of both stars (see \figname~\ref{fig:spec2}).
The measured absolute abundances for \bizarreone\ and \sunanalog\ are
$2.75$~dex and $2.25$~dex, respectively (Myles \etal, in prep),
resulting in the largest abundance difference ($\approx 0.5$~dex) seen
among all measured elements.

@@

We interpret the \Tcondens-dependent metal enhancement of \bizarreone\ relative
to \sunanalog\ and the very high surface \elem{Li} abundance ($A(\elem{Li}) =
2.75$) of \bizarreone\ as evidence for accretion of rocky (refractory-rich)
planetary material in \bizarreone.
In a multi-planet system, dynamical instabilities triggered by planet-planet
scattering\cite{} or encounters with a field star\cite{Malmberg:2011aa} can
lead to planet ejection or accretion.
If one of the two stars in a wide binary accreted more refractory-rich (rocky) planets
than the other, this may imprint a strong \Tcondens\-dependent
abundance differences like those observed in \bizarreone-\sunanalog.

@@

What triggered the planet engulfment in the two comoving stars
within the last Gyr remains unclear.
A possibility is that a fly-by interaction with a field star could have
triggered eccentricity excitation of outer planets around each star, which may
have propagated inward through planet-planet scattering, making inner rocky planets
to be accreted\cite{Malmberg:2011aa}.
If this is the case, there may be surviving highly eccentric giant planets that are
potentially detectable with future data releases of \gaia.
The two stars have not been included in any publicly released data from planet
search programs.
If both stars have accreted planetary material, it would be very interesting to
search for the existence and architectures of the planetary systems left
behind.

@@

@@

\subsection{Proximity in phase-space coordinates}

>>>>

% general description of the pair and measurements
\sunanalog\ and \bizarreone\ were identified as a candidate comoving star pair
in our recent search for comoving stars using the proper motions and parallaxes
from the {\it Tycho-Gaia Astrometric Solution} catalog (hereafter \tgas), a
component of the first data release of the astrometric space mission \gaia\cite{2016A&A...595A...2G}
(the astrometric measurements are listed in \tablename~\ref{tab:t1}).
The pair has also been previously recognized as a visual double star system
in Washington Double Star catalog\cite{2001AJ....122.3466M}.
The two stars have spectral types G0 and G2 both similar to the Sun (G2).
In a separate effort to study detailed chemical abundances of potential
planet-hosting stars, high resolution ($R=\lambda/\Delta\lambda\approx 70000$),
high signal-to-noise ratio (typically $>200$ per pixel in continuum) spectra of
both stars were obtained using the HIRES spectrograph on the Keck-I
telescope\cite{2016ApJS..225...32B}. 
The wavelength coverage is $5164$--$7799$~\AA.
The resulting measurements include elemental abundances for 15 chemical species
(C, N, O, Na, Mg, Al, Si, Ca, Ti, V, Cr, Mn, Fe, Ni, Y) as well as stellar
parameters and high precision ($\sigma\approx0.2$~\kms) radial velocities
(section~Table~\ref{tab:t1}).

% likely origin given its proximity in phase space
The projected separation between the pair is 1.9 arcmin\ ($\approx 0.01$~pc),
and the 3D separation is $\approx 0.6$~pc.
Although selected based only on their astrometry, the two stars
have identical radial velocities within uncertainties,
confirming that they are truly comoving (\ref{supp:dxdv}).
As the separation is smaller than the Jacobi radius (1.2~pc) in the solar
neighborhood for a 2~\msun\ binary system\cite{Jiang:2010aa}, \bizarreone\ and
\sunanalog\ are likely a bound wide binary system that formed coevally.

% coevality based on stellar parameters
% TODO reference supplement discussion on age here?
Their stellar parameters also indicate that they are coeval with an age of $\sim 4$~Gyr.
We use the distances (inferred from \gaia\ parallaxes), $V$-band magnitudes,
and $B-V$ colors to obtain bolometric luminosities of the two stars\cite{2003AJ....126..778V}.
We then combine the luminosities with $T_\mathrm{eff}$, \elemH{Fe}, \elemH{Si}
in order interpolate the age, mass, and radius of each star
in a grid of Yale-Yonsei model isochrones\cite{2013ApJ...776...87S}.
The best-fit isochrone ages of \bizarreone\ and \sunanalog\ are
$4.28_{-1.03}^{+1.11}$~Gyr and $4.00_{-1.56}^{+1.51}$~Gyr, respectively,
consistent with them being coeval (see Figure~\ref{fig:yyisofitting}).

% metallicity and abundance difference
However, one of the stars, \bizarreone, is significantly more metal rich than
the other by 0.2~dex in \feh\ ($\approx 60\%$; \figname~\ref{fig:abundances}).
Moreover, not all elements are equally enhanced: the abundances of \bizarreone\
show selective under-enhancement in \elem{C}, \elem{N}, \elem{O}, \elem{Na},
and \elem{Mn} relative to \elem{Fe} and other refractory elements.
As expected from their measured metallicity difference, the ratio of data and
model between the two stars show significant residuals of almost all metal line
features, largely dominated by \elem{Fe}.
However, for lines of elements that are not as enhanced in \bizarreone\, the
residuals are much smaller in amplitude demonstrating the validity of the
measured abundance differences (Figure~\ref{fig:spec1} and \ref{fig:spec2}).

% novelty
The observed difference in metallicity and abundances are genuinely rare and
unexpected.
None of the other four twin-like ($\Delta T_\mathrm{eff} \lesssim 100$~K) wide
binary pairs examined using spectra of similar quality and the same
consistent measurement pipeline shows discrepancies in abundances between the
stars at this level\cite{2016ApJS..225...32B} (Figure~\ref{fig:deltaXH}).
The maximum metallicity difference observed in the other pairs is $\Delta\feh =
0.03$~dex with $1\sigma$ statistical uncertainties in \feh\ of $0.01$~dex.
This is consistent with similar studies of metallicity and abundance
differences in wide binaries\cite{Gratton:2001aa,Desidera:2004aa}.
In fact, the abundance pattern of \bizarreone\ is a very unlikely one
for a star that is enhanced in \feh\ by $0.2$~dex relative to \sunanalog\
($\feh\approx 0$; \figname~\ref{fig:deltaXH}),
implying that a scenario involving Galactic chemical evolution is not plausible
even if we ignore a coeval star, \sunanalog, that we can compare with.
The difference of $\approx 0.2$~dex is also the largest among a handful of wide
binaries that do show metallicity and abundance differences with varying
significance\cite{Mack:2014aa,Mack:2016aa,Saffe:2015aa,
  Teske:2013aa,Teske:2015aa,Teske:2016aa,Teske:2016ab,Biazzo:2015aa,Ramirez:2015aa}
(Figure~\ref{fig:relabun_tcrank}).

% high Li
Both stars have high surface \elem{Li} abundances considering
their ages of $\sim 4$~Gyr.
The surface lithium abundance in a sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning. %TODO: REFERENCE?
The \elem{Li} doublet at $6707.6$~\AA, which is usually too weak to be
detectable at this age even for high signal-to-noise spectra, is clearly
visible in the spectra of both stars (\figname~\ref{fig:spec2}).
The measured absolute abundances for \bizarreone\ and \sunanalog\ are
$2.75$~dex and $2.25$~dex, respectively (Myles \etal, in prep),
resulting in the largest abundance difference ($\approx 0.5$~dex) seen
among all measured elements.

@@

We interpret the \Tcondens-dependent metal enhancement of \bizarreone\ relative
to \sunanalog\ and the very high surface \elem{Li} abundance ($A(\elem{Li}) =
2.75$) of \bizarreone\ as evidence for accretion of rocky (refractory-rich)
planetary material in \bizarreone.
In a multi-planet system, dynamical instabilities triggered by planet-planet
scattering\cite{} or encounters with a field star\cite{Malmberg:2011aa} can
lead to planet ejection or accretion.
If one of the two stars in a wide binary accreted more refractory-rich (rocky) planets
than the other, this may imprint a strong \Tcondens-dependent
abundance differences like those observed in \bizarreone-\sunanalog.

@@

What triggered the planet engulfment in the two comoving stars
within the last Gyr remains unclear.
A possibility is that a fly-by interaction with a field star could have
triggered eccentricity excitation of outer planets around each star, which may
have propagated inward through planet-planet scattering, making inner rocky planets
to be accreted\cite{Zakamska:2004aa,Malmberg:2011aa}.
If this is the case, there may be surviving highly eccentric giant planets that are
potentially detectable with future data releases of \gaia.
The two stars have not been included in any publicly released data from planet
search programs.
If both stars have accreted planetary material, it would be very interesting to
search for the existence and architectures of the planetary systems left
behind.

@@

\clearpage

@@

\subsection{Proximity in phase-space coordinates}
\label{supp:dxdv}
